CF Registry Analysis

### 2024-07-11 Christine Chan/CF Registry/cf_registry_survival_analysis.qmd
simplified incidence and prevalence using annualized data

  breaks = c(-Inf, 10, 18, 30, Inf),

@@

annual$age_group <- relevel(annual$age_group, ref = "[10,18)")

@@

# For each person, filter out all the years after first CFRD diagnosis
annual <- annual %>% filter(ReviewYear <= first_cfrd | is.na(first_cfrd))
# Remove those under 10
annual <- annual %>% filter(Age_YrEnd != 0, Age_YrEnd >= 10)
annual$age_group <- droplevels(annual$age_group)

@@

# CFRD incidence and prevalence

@@

# Incidence

## Full cohort

```{r}
#| message: false
og <- ymd("2003-01-01")
km_fit <- survfit(
  Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2,
    origin = as.numeric(og)
  ) ~ 1,
  data = annual, id = eDWID
)
# Plot
surv_p <- ggsurvplot(km_fit, data = annual, fun = "event")
surv_p$plot +
  scale_x_continuous(
    breaks = km_fit$time,
    labels = year(ymd("2003-01-01") + days(km_fit$time))
  ) +
  theme(axis.title.x = element_blank(), legend.position = "none") +
  ggtitle("Cumulative Incidence")
# Table
km_table <- tidy(km_fit)
km_table$time <- year(og + days(km_table$time))
gt(km_table) %>%
  cols_hide("std.error") %>%
  fmt_percent(columns = c("estimate", "conf.low", "conf.high")) %>%
  cols_merge(
    columns = c("estimate", "conf.low", "conf.high"),
    pattern = "{1} [{2} - {3}]"
  ) %>%
  cols_label(
    time = "Year", n.risk = "At Risk", n.event = "CFRD+",
    n.censor = "Censored", estimate = "Estimated CFRD- [95% CI]"
  ) %>%
  opt_row_striping() %>%
  opt_table_lines() %>%
  opt_stylize(style = 6, color = "gray")
```

```{r}
#| message: false
# Plot
km_age_fit <- survfit(
  Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2,
    origin = as.numeric(og)
  ) ~ age_group,
  data = annual, id = eDWID
)
surv_p <- ggsurvplot(km_age_fit, data = annual, fun = "event")
surv_p$plot +
  scale_x_continuous(
    breaks = km_age_fit$time,
    labels = year(ymd("2003-01-01") + days(km_age_fit$time))
  ) +
  scale_color_discrete(name = "Age Group", labels = c("10-17", "18-29", "30+")) +
  theme(axis.title.x = element_blank(), legend.title = element_blank()) +
  ggtitle("Cumulative Incidence by Age Group")
# Table
km_table <- tidy(km_age_fit)
km_table$time <- year(og + days(km_table$time))
km_table <- split.data.frame(km_table, km_table$strata)

invisible(lapply(km_table, function(km) {
  title <- km$strata[1]
  title <- sub("age_group=", "Ages ", title)
  gt(km) %>%
    cols_hide(c("std.error", "strata")) %>%
    fmt_percent(columns = c("estimate", "conf.low", "conf.high")) %>%
    cols_merge(
      columns = c("estimate", "conf.low", "conf.high"),
      pattern = "{1} [{2} - {3}]"
    ) %>%
    cols_label(
      time = "Year", n.risk = "At Risk", n.event = "CFRD+",
      n.censor = "Censored", estimate = "Estimated CFRD- [95% CI]"
    ) %>%
    opt_row_striping() %>%
    opt_table_lines() %>%
    opt_stylize(style = 6, color = "gray") %>%
    tab_caption(title)
}))
```

@@

```{r}
# How many OGTTs are non-CFRD people getting 2016-2018
ogtts <- annual %>%
  filter(ReviewYear %in% 2016:2018, CFRD == "CFRD-") %>%
  group_by(eDWID) %>%
  summarise(
    `Age Group` = last(na.omit(age_group)),
    `OGTTs` = sum(!is.na(A_twohour_bloodglucose) | !is.na(A_ogttfast_bloodglucose))
  )
# Add demographics
ogtts$eDWID <- as.numeric(ogtts$eDWID)
ogtts <- left_join(ogtts, demogs, by = join_by(eDWID))
label(ogtts$eDWID) <- "ID"
# Format OGTT counts
ogtts$OGTTs <- factor(ogtts$OGTTs,
  levels = 0:3, labels = c(
    "No OGTT", "At least 1 OGTT", "At least 1 OGTT", "At least 1 OGTT"
  )
)
# Compare no OGTT to at least 1 OGTT
ogtts %>% 
  select(-eDWID) %>%
  tbl_summary(by=OGTTs,missing_text = "Missing") %>%
  add_p()
```

@@

```{r}
# Formula for base model
xvars <- c(
  "age_group", "Gender", "Race", "Hispanicrace", "bmi_perc_cat", "MutClass",
  "A_IsOnEnzymes", "tube_feeding", "liver_disease", "ppFEV1_cat",
  "NumPulmExacerbation", "A_corticosteroids1"
)
y <- "Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2)"
f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
cox_res <- coxph(f, data = annual, id = eDWID)
# Plot survival
# ggsurvplot(survfit(cox_res),
#   data = s, risk.table = T, cumevents = T, fun = "event",
#   xlim = c(2003, 2018), break.x.by = 1
# )
# Forest plot and coefficients table
cox_res %>%
  tbl_regression() %>%
  add_n(location = "level") %>%
  add_nevent(location = "level")
ggcoef_model(cox_res,
  show_p_values = F, add_reference_rows = F,
  signif_stars = F, colour = F,
)
# Diagnostics
test.ph <- data.frame(cox.zph(cox_res)$table)
test.ph <- test.ph[rownames(test.ph) != "GLOBAL", ]
# Find which variables technically violate PH assumption
time_varying <- rownames(test.ph)[test.ph$p < 0.05]
```

>>>>

  breaks = c(-Inf, 10, 19, 31, Inf),

@@

@@

@@

# CFRD prevalence

## By year

```{r}
annual %>%
  group_by(ReviewYear) %>%
  summarise(Prevalence = mean(CFRD == "CFRD+") * 100) %>%
  kable(., digits = 2)
```

## By year and age group

```{r}
annual %>%
  group_by(ReviewYear, age_group) %>%
  summarise(Prevalence = mean(CFRD == "CFRD+") * 100, .groups = "drop") %>%
  pivot_wider(names_from = age_group, values_from = Prevalence) %>%
  kable(., digits = 2)
```

# CFRD incidence

```{r}
# For each person, filter out all the years after first CFRD diagnosis
annual <- annual %>% filter(ReviewYear <= first_cfrd | is.na(first_cfrd))
# For each year, pull everyone with and without CFRD that year
year_inc <- function(df = annual, year, py = 1000) {
  # Filter and count
  inc <- df %>% filter(ReviewYear == year)
  inc <- inc %>%
    group_by(CFRD) %>%
    count(age_group, .drop = F)
  inc <- inc[!is.na(inc$age_group), ]
  # Calculations
  new_cases <- sum(inc$n[inc$CFRD == "CFRD+"])
  total_at_risk <- sum(inc$n)
  total_inc <- new_cases / (total_at_risk / py)

  new_cases_10 <- sum(inc$n[inc$CFRD == "CFRD+" & inc$age_group == "[-Inf,10)"])
  at_risk_10 <- sum(inc$n[inc$age_group == "[-Inf,10)"])
  inc_10 <- new_cases_10 / (at_risk_10 / py)

  new_cases_10_18 <- sum(inc$n[inc$CFRD == "CFRD+" & inc$age_group == "[10,19)"])
  at_risk_10_18 <- sum(inc$n[inc$age_group == "[10,19)"])
  inc_10_18 <- new_cases_10_18 / (at_risk_10_18 / py)

  new_cases_19_30 <- sum(inc$n[inc$CFRD == "CFRD+" & inc$age_group == "[19,31)"])
  at_risk_19_30 <- sum(inc$n[inc$age_group == "[19,31)"])
  inc_19_30 <- new_cases_19_30 / (at_risk_19_30 / py)

  new_cases_30 <- sum(inc$n[inc$CFRD == "CFRD+" & inc$age_group == "[31, Inf)"])
  at_risk_30 <- sum(inc$n[inc$age_group == "[31, Inf)"])
  inc_30 <- new_cases_30 / (at_risk_30 / py)

  # Nice results
  list(
    `Year` = year,
    `Total New Cases` = new_cases,
    `Total At Risk` = total_at_risk,
    `Total Incidence` = total_inc,
    `Total New Cases Age < 10` = new_cases_10,
    `Total At Risk Age < 10` = at_risk_10,
    `Total Incidence Age < 10` = inc_10,
    `Total New Cases Age 10-18` = new_cases_10_18,
    `Total At Risk Age 10-18` = at_risk_10_18,
    `Total Incidence Age 10-18` = inc_10_18,
    `Total New Cases Age 19-30` = new_cases_19_30,
    `Total At Risk Age 19-30` = at_risk_19_30,
    `Total Incidence Age 19-30` = inc_19_30,
    `Total New Cases Age >= 31` = new_cases_30,
    `Total At Risk Age >= 31` = at_risk_30,
    `Total Incidence Age >= 31` = inc_30
  )
}
incidence <- lapply(2003:2018, function(y) {
  year_inc(year = y)
})
incidence <- data.frame(do.call(rbind, incidence), check.names = F)
incidence <- data.frame(lapply(incidence, as.numeric), check.names = F)
kable(incidence, digits = 2)
```

@@

@@

```{r}
# How many OGTTs are non-CFRD people getting 2016-2018
ogtts <- annual %>%
  filter(ReviewYear %in% 2016:2018, CFRD == "CFRD-") %>%
  group_by(eDWID) %>%
  summarise(
    `Age Group` = last(na.omit(age_group)),
    `OGTTs` = sum(!is.na(A_twohour_bloodglucose) | !is.na(A_ogttfast_bloodglucose))
  )
# Add demographics
ogtts$eDWID <- as.numeric(ogtts$eDWID)
ogtts <- left_join(ogtts, demogs, by = join_by(eDWID))
label(ogtts$eDWID) <- "ID"
# Format OGTT counts
ogtts$OGTTs <- factor(ogtts$OGTTs,
  levels = 0:3, labels = c(
    "No OGTT", "At least 1 OGTT", "At least 1 OGTT", "At least 1 OGTT"
  )
)
# Compare no OGTT to at least 1 OGTT
ogtts %>%
  select(-eDWID) %>%
  tbl_summary(by = OGTTs, missing_text = "Missing") %>%
  add_p()
```

@@

```{r}
# Remove those under 10
annual$age_group <- relevel(annual$age_group, ref = "[10,19)")
annual <- annual %>% filter(Age_YrEnd != 0, Age_YrEnd >= 10)
annual$age_group <- droplevels(annual$age_group)
# Origin for survival analyses
og <- ymd("2003-01-01")
# Formula for base model
xvars <- c(
  "age_group", "Gender", "Race", "Hispanicrace", "bmi_perc_cat", "MutClass",
  "A_IsOnEnzymes", "tube_feeding", "liver_disease", "ppFEV1_cat",
  "NumPulmExacerbation", "A_corticosteroids1"
)
y <- "Surv(as.numeric(tstart), as.numeric(tstop), as.numeric(CFRD) == 2)"
f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
cox_res <- coxph(f, data = annual, id = eDWID)
# Plot survival
# ggsurvplot(survfit(cox_res),
#   data = s, risk.table = T, cumevents = T, fun = "event",
#   xlim = c(2003, 2018), break.x.by = 1
# )
# Forest plot and coefficients table
cox_res %>%
  tbl_regression() %>%
  add_n(location = "level") %>%
  add_nevent(location = "level")
ggcoef_model(cox_res,
  show_p_values = F, add_reference_rows = F,
  signif_stars = F, colour = F,
)
# Diagnostics
test.ph <- data.frame(cox.zph(cox_res)$table)
test.ph <- test.ph[rownames(test.ph) != "GLOBAL", ]
# Find which variables technically violate PH assumption
time_varying <- rownames(test.ph)[test.ph$p < 0.05]
```
